\(ler\) : LVK (LIGO-Virgo-KAGRA collaboration) Event (compact-binary mergers) Rate calculator and simulator

### 2024-06-26 paper/paper.tex
mac

\(ler\) is a statistical-based Python package specifically designed for
computing detectable rates of both lensed and unlensed gravitational
wave (GW) events, catering to the requirements of the LIGO-Virgo-KAGRA
Scientific Collaboration and astrophysics research scholars. The core
functionality of \(ler\) intricately hinges upon the interplay of
various components which include sampling the properties of
compact-binary sources, lens galaxies characteristics, solving lens
equations to derive properties of resultant images, and computing
detectable GW rates. This comprehensive functionality builds on the
leveraging of array operations and linear algebra from the \emph{numpy}
(NumPy Community 2022) library, enhanced by interpolation methods from
\emph{scipy} (Virtanen et al. 2020) and Python's \emph{multiprocessing}
capabilities. Efficiency is further boosted by the \emph{numba} (Lam,
Pitrou, and Seibert 2022) library's Just-In-Time (\emph{njit})
compilation, optimizing extensive numerical computations and employing
the inverse transform sampling method to replace more cumbersome
rejection sampling. The modular design of \(ler\) not only optimizes
speed and functionality but also ensures adaptability and upgradability,
supporting the integration of additional statistics as research evolves.
Currently, \(ler\) is an important tool in generating simulated
gravitational wave (GW) events---both lensed and unlensed---and provides
astrophysically accurate distributions of event-related parameters for
both detectable and non-detectable events. This functionality aids in
event validation and enhances the forecasting of detection capabilities
across various GW detectors to study such events. The architecture of
the \(ler\) API facilitates seamless compatibility with other software
packages, allowing researchers to integrate and utilize its
functionalities based on specific scientific requirements.

@@

\begin{itemize}
\tightlist
\item
  Sampling Gravitational Wave (GW) Source Properties:

@@

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-Ashton2019}
Ashton, Gregory, Moritz Hübner, Paul D. Lasky, Colm Talbot, Kendall
Ackley, Sylvia Biscoveanu, Qi Chu, et al. 2019. {``Bilby: A
User-Friendly Bayesian Inference Library for Gravitational-Wave
Astronomy.''} \emph{The Astrophysical Journal Supplement Series} 241
(2): 27. \url{https://doi.org/10.3847/1538-4365/ab06fc}.

@@

>>>>

\(ler\) is a statistics-based Python package specifically designed for
computing detectable rates of both lensed and unlensed GW events,
catering to the requirements of the LIGO-Virgo-KAGRA Scientific
Collaboration (The LIGO Scientific Collaboration et al. 2015) (Acernese
et al. 2014) (Akutsu et al. 2020) and astrophysics research scholars.
The core functionality of \(ler\) intricately hinges upon the interplay
of various components which include sampling the properties of
compact-binary sources, lens galaxies characteristics, solving lens
equations to derive properties of resultant images, and computing
detectable GW rates. This comprehensive functionality builds on the
leveraging of array operations and linear algebra from the \emph{numpy}
(NumPy Community 2022) library, enhanced by interpolation methods from
\emph{scipy} (Virtanen et al. 2020) and Python's \emph{multiprocessing}
capabilities. Efficiency is further boosted by the \emph{numba} (Lam,
Pitrou, and Seibert 2022) library's Just-In-Time (\emph{njit})
compilation, optimizing extensive numerical computations and employing
the inverse transform sampling method to replace more cumbersome
rejection sampling. The modular design of \(ler\) not only optimizes
speed and functionality but also ensures adaptability and upgradability,
supporting the integration of additional statistics as research evolves.
Currently, \(ler\) is an important tool in generating simulated GW
events---both lensed and unlensed---and provides astrophysically
accurate distributions of event-related parameters for both detectable
and non-detectable events. This functionality aids in event validation
and enhances the forecasting of detection capabilities across various GW
detectors to study such events. The architecture of the \(ler\) API
facilitates seamless compatibility with other software packages,
allowing researchers to integrate and utilize its functionalities based
on specific scientific requirements.

@@

\begin{itemize}
\tightlist
\item
  Sampling GW Source Properties:

@@

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-VIRGO2015}
Acernese, F, M Agathos, K Agatsuma, D Aisa, N Allemandou, A Allocca, J
Amarni, et al. 2014. {``Advanced Virgo: A Second-Generation
Interferometric Gravitational Wave Detector.''} \emph{Classical and
Quantum Gravity} 32 (2): 024001.
\url{https://doi.org/10.1088/0264-9381/32/2/024001}.

\bibitem[\citeproctext]{ref-KAGRA2021}
Akutsu, T, M Ando, K Arai, Y Arai, S Araki, A Araya, N Aritomi, et al.
2020. {``{Overview of KAGRA: Detector design and construction
history}.''} \emph{Progress of Theoretical and Experimental Physics}
2021 (5): 05A101. \url{https://doi.org/10.1093/ptep/ptaa125}.

\bibitem[\citeproctext]{ref-Ashton2019}
Ashton, Gregory, Moritz Hübner, Paul D. Lasky, Colm Talbot, Kendall
Ackley, Sylvia Biscoveanu, Qi Chu, et al. 2019. {``Bilby: A
User-Friendly Bayesian Inference Library for Gravitational-Wave
Astronomy.''} \emph{The Astrophysical Journal Supplement Series} 241
(2): 27. \url{https://doi.org/10.3847/1538-4365/ab06fc}.

@@

\bibitem[\citeproctext]{ref-LIGO2015}
The LIGO Scientific Collaboration, J Aasi, B P Abbott, R Abbott, T
Abbott, M R Abernathy, K Ackley, et al. 2015. {``Advanced LIGO.''}
\emph{Classical and Quantum Gravity} 32 (7): 074001.
\url{https://doi.org/10.1088/0264-9381/32/7/074001}.
